Molecular dynamics simulation of synchronization in driven particles

### 2021-01-19 manuscript/manuscript_ajp_2020.tex
worked on exercises

  Reasonable experimental values for
  a colloid are $D = 1 \mu m$ and $v \sim 1 \mu m /s$.
  Viscosity of water decreases with temperature
  from $1.3 \times 10^{-3}$ Pa-s at $10^{\circ}$ C to 
  from $0.31 \times 10^{-3}$ Pa-s at $90^{\circ}$ C,
  making $\eta \sim 10^{-3}$ Pa-s,
  where 1 Pascal is $1$ N/m$^2$.
  Likewise the density $\rho$ varies with temperature
  and $\rho \sim 10^3$ kg/cm$^3$ is a reasonable approximation.
  \cite{asce}

@@

  {\it this isn't very challenging, but I've see undergraduate students have a ``wow'' factor when they work this out for the first time.}

  \item {\bf Integration Methods}
  Our equation of motion provides
  a direct calculation for particle velocity.
  To calculate the position of the particle we
  integrate the equation of motion using
  the standard definition of velocity
  $\vec{v}_i = d\vec{r}_i/dt$.
  %
  In our model we use the
  Euler method to solve this equation.
  Generally when solving ordinary differential equations,
  the Euler method is effective for solving linear equations
  of the form
  $dy/dt = f(t,y(t))$ with initial condition $y(t_0) = y_0$.
  The solution is calculated algorithmically
  by stepping in time through $n$ integer steps
  $t_n = t_0 + n \Delta t$.
  At each subsequent step the new
  value for $y$ is calculated as a map solution using
  discrete times 
  $y_{n+1} = y_n + f(t_n,y_n)$.
  The Euler method can be applied to non-linear
  equations effectively if the time step $\Delta t$
  is kept sufficiently small.  
  \cite{Newman}

  Generally molecular dynamics algorithms
  are solved with higher order methods
  such as the Verlet method or Runge-Kutta methods.

  and demonstrate how the Verlet method
  simplifies to the Euler method when $\vec{a}=0$.

  %FOLLOW UP WITH THIS!
  Because we are performing molecular dynamics
  for a single particle on a smooth potential energy landscape,
  we use a large simulation time step of 
  $\Delta t = 0.01$.
  In Sec.~\ref{sec:sync},
  we decrease the time step
  to treat multiple particle interactions.

  \label{ex:euler}

  We invite interested students to repeat this
  simulation for any initial particle position %of $y$
  and a range of simulation constants
  to demonstrate that synchronization in this model
  does not depend on a particular set of conditions.
  %but is a direct consequence of the simplicity of the model.

\item {\bf Visualizing the motion}
  Drawing the contour

@@

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference
%Fundamental models
%of synchronization 
%can %include many coupled oscillators
%or
%focus on 
A single particle driven across a periodic potential landscape 
%Single particle studies are useful
to understand synchronization
in the absence of collective effects
due to interactions between particles.
%A single particle particle oscillator
%within a potential well is 
%much like a skateboarder in a half-pipe or
%a child on a swing
%moving back and forth. % through equilibrium.  % \cite{}.
The relationship between the driving force and
landscape can change the pattern of motion.  
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning \cite{Reichhardt2017}.  
%and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.
Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments and simulations of 
driven colloids on a
optical periodic landscape.
Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE

@@

\bibitem{supp1} See Figure1.mp4 in appropriate %\url{}

\bibitem{asce}
  What is the method for citing non-fundamental constants?
    (1) https://ascelibrary.org/doi/pdf/10.1061/9780784408230.ap02
    (2) IAPWS 2008 - http://www.iapws.org/relguide/viscosity.html
  %[citation for these values better than wikipedia] \cite{} %CONFIRM!
  %density $\rho = 0.9982 g/cm^3$

>>>>

  To first order 
  approximate experimental values for
  a colloid are $D = 1 \mu m$ and $v \sim 1 \mu m /s$.
  Viscosity of water decreases with temperature
  from $1.3 \times 10^{-3}$ Pa-s at $10^{\circ}$ C to 
  from $0.31 \times 10^{-3}$ Pa-s at $90^{\circ}$ C,
  making $\eta \sim 10^{-3}$ Pa-s,
  where 1 Pascal is $1$ N/m$^2$.
  Likewise the density $\rho$ varies with temperature
  but  $\rho \sim 10^3$ kg/cm$^3$ is a reasonable
  first order approximation.
  \cite{asce}

@@

  {\it this isn't very challenging, but I've see undergraduate students have a ``wow'' factor when they work this out for the first time.  This is closely related to the next exercise which is also primarily analytic rather than computational}

  \item {\bf Integration Methods}
    \label{ex:euler}
    Our equation of motion provides
  a direct calculation for particle velocity.
  To calculate the position of the particle we
  integrate the equation of motion using
  the standard definition of velocity
  $\vec{v}_i = d\vec{r}_i/dt$ 
  via the 
  %In our model we use the
  Euler method. % to solve this equation.
  Generally when solving ordinary differential equations,
  the Euler method is effective for solving linear equations
  of the form
  $dy/dt = f(t,y(t))$ with initial condition $y(t_0) = y_0$.
  The solution is calculated algorithmically
  by stepping in time through $n$ integer steps
  $t_n = t_0 + n \Delta t$.
  At each subsequent step the new
  value for $y$ is calculated as a map solution using
  discrete times 
  $y_{n+1} = y_n + f(t_n,y_n)$.
  The Euler method can be applied to non-linear
  equations effectively if the time step $\Delta t$
  is kept sufficiently small \cite{Newman}.
  Apply the Euler method to our equation
  of motion to solve for the position of a particle
  $y_n$ at the $nth$ timestep.

  {\it invite/guide students to play with the timestep and study the results}
  %FOLLOW UP WITH THIS!
  %Because we are performing molecular dynamics
  %for a single particle on a smooth potential energy landscape,
  %we use a large simulation time step of 
  %$\Delta t = 0.01$.
  %In Sec.~\ref{sec:sync},
  %we decrease the time step
  %to treat multiple particle interactions.

  Often molecular dynamics algorithms 
  are solved with higher order methods
  such as the Verlet method or Runge-Kutta methods.
  [{\it explain Verlet and/or RK4...}]
  Demonstrate how the Verlet method
  simplifies to the Euler method when
  the particle acceleration $\vec{a}=0$.

  {\it invite/guide students to play with the timestep and study the results in comparison to Euler method}

  %We invite interested students to repeat this
  %simulation for any initial particle position %of $y$
  %and a range of simulation constants
  %to demonstrate that synchronization in this model
  %does not depend on a particular set of conditions.
  %but is a direct consequence of the simplicity of the model.

@@

  Drawing the contour

  Phase plots...

@@

\section{Conclusion}
\label{sec:conclusion}	% You can label sections for reference
%Fundamental models
%of synchronization 
%can %include many coupled oscillators
%or
%focus on 
A single particle driven across a periodic potential landscape 
%Single particle studies are useful
is useful 
to study synchronization
in the absence of collective effects of particle interactions.
%due to interactions between particles.
%A single particle particle oscillator
%within a potential well is 
%much like a skateboarder in a half-pipe or
%a child on a swing
%moving back and forth. % through equilibrium.  % \cite{}.
The relationship between the driving force and
landscape can change the pattern of motion.  
With a constant or 
dc drive the landscape modulates 
the particle velocity, below some threshold  
the dc force is not strong enough to push the particle
across a potential maximum so the average velocity is zero,
a phenomena referred to as pinning \cite{Reichhardt2017}.  
%and dissipating energy from the system at all driving forces.
A sinusoidal or ac drive creates mode-locking,
where the average particle velocity
is fixed for a range of dc drive forces \cite{Reichhardt2015}.
Our simulations reproduce results presented in 
Juniper {\it et al.} \cite{Juniper2015, Juniper2018}
which demonstrated
mode locking in
experiments and simulations of 
driven colloids on a
optical periodic landscape.
Due to the relative ease of
manipulation and imagining, 
colloids are used as experimental models
for systems relatively hard to access and visualize. 
%SAME SENTENCE TWICE

@@

      \bibitem{supp1} Supplementary videos coming soon.
        %See Figure1.mp4 in appropriate %\url{}

\bibitem{asce}
  %Non-fundamental constants for water viscosity are available
  %in several databases including 
  %  (1) https://ascelibrary.org/doi/pdf/10.1061/9780784408230.ap02
  IAPWS R12-08, 
  Release on the IAPWS Formulation 2008 for the Viscosity of Ordinary Water Substance, 
  September 2008, 
  %IAPWS 2008 -
  http://www.iapws.org/relguide/viscosity.html
  %[citation for these values better than wikipedia] \cite{} %CONFIRM!
  %density $\rho = 0.9982 g/cm^3$
